---
title: "Practical Machine Learning Course Project"
author: "Eduardo V. Kuri"
output: html_document
---

## Retrieval of Data

> First off, we need to obtain our data and install the caret library. We will do that in the following code snippet:

```{r setup}

require(caret)
require(ggplot2)
require(randomForest)


if(!exists("trainraw")){
  
  trainraw <- read.csv("~/UNAM 2020-24/DATSCI/MachineLearning/pml-training.csv")

}

if(!exists("testraw")){
  
  testraw <- read.csv("~/UNAM 2020-24/DATSCI/MachineLearning/pml-testing.csv")

}
```

```{r expana}

print(c(dim(testraw)[1], dim(trainraw)[1]))

per <- 100*dim(trainraw)[1]/(dim(testraw)[1]+dim(trainraw)[1])

print(paste("Percentage of training data:", round(per,4),"%", ", Percentage of testing data:", round(100-per,4),"%", ", Number of variables: ",dim(trainraw)[2]))

completecases <- complete.cases(trainraw)
print(length(completecases[completecases==TRUE]))


#Validation set


# Juntar modelos
```

> There seems to be a lot of information, and no guarantee that information will be of any use. We then clean the columns with NAs, in order to reduce the computational burden:

```{r cleanup}
trainraw <- trainraw[, colSums(is.na(trainraw)) == 0]
testraw <- testraw[, colSums(is.na(testraw))==0]
classe <- trainraw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainraw))
trainraw <- trainraw[, !trainRemove]
train <- trainraw[, sapply(trainraw, is.numeric)]
train$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testraw))
testraw <- testraw[, !testRemove]
test <- testraw[, sapply(testraw, is.numeric)]




belt <- grep("[Bb]elt", names(train))
farm <- grep("[Ff]orearm", names(train))
arm <- grep("[Aa]rm", names(train))
dbell <- grep("[Dd]umbbell", names(train))
classe0 <- grep("classe", names(train))


belttrain <- train[,c(belt, classe0)] 
farmtrain <- train[,c(farm, classe0)]
armtrain <- train[,c(arm,classe0)]
dbelltrain <- train[,c(dbell, classe0)]

#Validation set


# Juntar modelos
```

## Slice into Train and Validation sets

```{r slice}

set.seed(123455)

intrain <- createDataPartition(train$classe, p = 0.75, list=F)

training <- train[intrain, ]
val <- train[-intrain, ]




  # Juntar modelos
```

## Training

```{r training}
# instruct train to use 3-fold CV to select optimal tuning parameters
fitctrl <- trainControl(method="cv", number=3, verboseIter=F)
mod0 <- train(classe ~ ., data=training, method="rf", trControl = fitctrl)

mod0$finalmodel
```

## Prediction and Results

```{r pred}

pred1 <- predict(mod0, testing[,-length(names(testing))])

print(pred1)


```

## Conclusions

## References

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.
